Optimization of String Rewriting Operations for 3D Fractal Generation with Genetic Algorithms

### 2018-07-04 Optimization-of-String-Rewriting-Operations-for-3D-Fractal-Generation-with-Genetic-Algorithms/paper.tex
Some remarks ware fixed.

\usepackage{graphicx}
\usepackage{placeins}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amssymb}

@@

String rewriting system (SRS) is a substitution system in which the rules does not contain nonterminal symbols. Each terminal symbol can appear on the LHS of the rule and on the RHS of the rule. With such definition SRS are useful for generation of infinite structures and they are mainly used for the generation of some fractals (for example: Box Fractal, Cantor Dust, Cantor Square Fractal and Sierpinski Carpet). Most of the fractals are generated with binary rules (black and white colors), but in the Hodkinson's research it is clearly shown that limit of the colors can be higher.

@@

This study extends the idea from 2D space into 3D space. Voxels are used instead of pixels. Square matrix of 3x3 is replaced with cube 3x3x3. The paper is organized as follows: Section \ref{Introduction} introduces the problem; Section \ref{Model and Optimization} presents a model and optimization approach; Section \ref{Experiments and Results} gives experiment details and the results are shown; Section \ref{Conclusions} concludes and some further ideas for research are pointed.

@@

A finite 3D space is represented as three-dimensional array of voxels in the model. Each voxel is encoded with a single long integer value, which consists of the RGB information for the color of the voxel. In this model 32 different transition 3x3x3 matrices are selected, each corresponding to the one of the 32 colors. The steps for fractal generation are given in Algorithm \ref{algorithm01}.

@@

The individuals in the genetic algorithm are presented as 32 transition matrices each with size 3x3x3. The order of the matrices shows which transition will be applied according to used colors order. The steps of the optimization are illustrated in Algorithm \ref{algorithm02}.

@@

Tournament selection with arity of two is proposed. Uniform crossover is taken as the most symmetric crossover. From each parent randomly are taken 50\% of the elements. Random change of a color is used for the mutation operator. The evaluation is done by application of each 32 transition matrices with the certain level of recursion depth. Euclidean distance between generated fractal and target shape is used as fitness function. The initial fractal is a single cube painted with the color with the first index in the order of the colors. Genetic algorithms are global heuristics optimization and this means that it does not matter what kind of color the initial voxel will have.

@@

All experiments are done in Java programming language with Apache Genetic Algorithms Framework \cite{apache01} as an optimization approach and Printing in 3D - JavaSCAD \cite{printing01} as 3D visualization tool. The source code used for all experiments is an open source and can be found at \cite{balabanov01}. Increasing the space dimensions from 2D to 3D greatly increases the complexity of the optimization process. Because of the increased complexity less ambitious shape than the Greek letter $\pi$ is taken, in our case it is a voxelized sphere (Fig. \ref{fig:pic02}). For visualization purposes the first color in the set of colors is accepted as transparent.

>>>>

\usepackage{graphicx}
\usepackage{placeins}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amssymb}
\usepackage{listings}

@@

String rewriting system (SRS) is a substitution system in which the rules does not contain nonterminal symbols. Each terminal symbol can appear on the left hand side (LHS) of the rule and on the right hand side (RHS) of the rule. With such definition SRS are useful for generation of infinite structures and they are mainly used for the generation of some fractals (for example: Box Fractal, Cantor Dust, Cantor Square Fractal and Sierpinski Carpet). Most of the fractals are generated with binary rules (black and white colors), but in the Hodkinson's research it is clearly shown that limit of the colors can be higher.

@@

This study extends the idea from 2D space into 3D space, which is the main contribution of the authors. Voxels are used instead of pixels. Square matrix of 3x3 is replaced with cube 3x3x3. The paper is organized as follows: Section \ref{Introduction} introduces the problem; Section \ref{Model and Optimization} presents a model and optimization approach; Section \ref{Experiments and Results} gives experiment details and the results are shown; Section \ref{Conclusions} concludes and some further ideas for research are pointed.

@@

A finite 3D space is represented as three-dimensional array of voxels in the model. Each voxel is encoded with a single long integer value, which consists of the RGB (red, green, blue) information for the color of the voxel. In this model 32 different transition 3x3x3 matrices are selected, each corresponding to the one of the 32 colors. The steps for fractal generation are given in Algorithm \ref{algorithm01}.

@@

The individuals in the genetic algorithm are presented as 32 transition matrices each with size 3x3x3 (Listing \ref{listing03} in \cite{balabanov01}). Long integer numbers are used for RGB color representation for all 27 cells of the transition matrix.

\begin{lstlisting}[language=Java, caption=Chromosomes Encoding, label=listing03]
List<Chromosome> list = new LinkedList<Chromosome>();

for (int i = 0; i < 37; i++) {
	Long cells[] = new Long[COLORS.length * 3*3*3];

	for (int j = 0; j < cells.length; j++) {
		cells[j] = 
		COLORS[PRNG.nextInt(COLORS.length)];
	}

	list.add(new TransitionsChromosome(cells, 
	RECURSION_DEPTH, COLORS, target, start));
}
\end{lstlisting}

The order of the matrices shows which transition will be applied according to used colors order. The steps of the optimization are illustrated in Algorithm \ref{algorithm02}.

@@

Parents in the population are appointed by tournament selection with arity of two (Listing \ref{listing01} in \cite{balabanov01}).

\begin{lstlisting}[language=Java, caption=Genetic Algorithm Parameters, label=listing01]
GeneticAlgorithm optimizer = new GeneticAlgorithm(
new UniformCrossover<TransitionsChromosome>(0.5), 0.9, 
new TransitionsMutation(COLORS), 0.01, 
new TournamentSelection(2));
\end{lstlisting}

Uniform crossover is taken as the most symmetric crossover (Listing \ref{listing01} in \cite{balabanov01}). From each parent randomly are taken 50\% of the elements. Random change of a color is used for the mutation operator (Listing \ref{listing02} in \cite{balabanov01}).

\begin{lstlisting}[language=Java, caption=Single Voxel Color Mutation, label=listing02]
List<Long> values = new 
ArrayList<Long>(((TransitionsChromosome) original).
getRepresentation());

values.set(PRNG.nextInt(values.size()), 
colors[PRNG.nextInt(colors.length)]);
\end{lstlisting}

The evaluation is done by application of each 32 transition matrices with the certain level of recursion depth. Euclidean distance between generated fractal and target shape is used as fitness function. The initial fractal is a single cube painted with the color with the first index in the order of the colors. Genetic algorithms are global heuristics optimization and this means that it does not matter what kind of color the initial voxel will have.

@@

All experiments are done in Java programming language with Apache Genetic Algorithms Framework \cite{apache01} as an optimization approach and Printing in 3D - JavaSCAD \cite{printing01} as 3D visualization tool. The source code used for all experiments is an open source and can be found at \cite{balabanov01} for the interested readers who would like to reproduce the experiments and results. Increasing the space dimensions from 2D to 3D greatly increases the complexity of the optimization process. Because of the increased complexity less ambitious shape than the Greek letter $\pi$ is taken, in our case it is a voxelized sphere (Fig. \ref{fig:pic02}). For visualization purposes the first color in the set of colors is accepted as transparent.
